Assessing Ranking Fidelity of Competition Structures via Weighted Mutual Information

### 2025-10-21 CMSAC_presentation/CMSAC_reveal_js.qmd
Trimmed Slides

---
title: "Assessing Ranking Fidelity of Competition Structures via Weighted Mutual Information"
author: "Zach Culp, Josie Peterburs, Dr. Gregory J. Matthews, Dr. Ryan P.A. McShane"
format: 
  revealjs: 
    df-print: paged
editor: visual
---

@@

-   A limitation of mutual information is that it does not distinguish between the direction of information
-   In other words, given equal probabilities, a ranking of $(1,2,3,4)$ is equivalent to $(4,3,2,1)$
-   Because of this, we proposed a new weighting function $w(r, \hat{r})$\
    $$ wI_m(r,\hat{r}) = \sum_r \sum_{\hat{r}} w(r,\hat{r}) * p(r, \hat{r}) \log{\frac{p(r,\hat{r})}{p(r)p(\hat{r})}} $$

@@

## Weighted Kendall

-   If $\hat{r} = (2,1,3,4)$, then:

$$\sum_{i\neq j} W_{ij} \text{ sgn}(i-j) \text{ sgn}(R_i - R_j) = 10$$

-   So, $\tau_w$ for this example would be $\frac{10}{15-3} = 0.833$

@@

## Comparing Weighted Kendall's Tau to Squared Error

```{r}
# Get the filtered data
table_data <- rhat_df %>%    
  filter(k == 4) %>%   
  select(-k) %>%   
  arrange(desc(weighted_kendall)) %>%    
  rename("Weighted Kendall" = weighted_kendall, "Squared Error" = squared_error)

# Create side-by-side tables
left_html <- left_table %>% 
  kable(format = "html", row.names = FALSE) %>% 
  kableExtra::kable_styling(font_size = 20)
right_html <- right_table %>% 
  kable(format = "html", row.names = FALSE) %>% 
  kableExtra::kable_styling(font_size = 20)

## Normalizing Mutual Information

-   To make comparisons across different values of $n$, we normalize by the joint entropy

-   So, the final formula for the metric is:

$$\frac{wI_m(r,\hat{r})}{H(r,\hat{r})} = \frac{\sum_r \sum_{\hat{r}} w(r,\hat{r}) * p(r, \hat{r}) \log{\frac{p(r,\hat{r})}{p(r)p(\hat{r})}}}{\log_2((n!)^2)}$$

-   Assuming $r$ and $\hat{r}$ are independent, joint entropy can be written as $H(r, \hat{r}) = H(r) + H(\hat{r})$

>>>>

---
title: "Assessing Ranking Fidelity of Competition Structures via Weighted Mutual Information"
author: "Zach Culp, Josie Peterburs, Dr. Gregory J. Matthews, Dr. Ryan P.A. McShane"
format: 
  revealjs: 
    html-math-method: 
      method: mathjax
      url: "https://cdn.jsdelivr.net/npm/mathjax@3/es5/tex-mml-chtml.js"
    df-print: paged
editor: visual

---

@@

-   A limitation of mutual information is that it does not distinguish between the direction of information
-   In other words, given equal probabilities, a ranking of $(1,2,3,4)$ is equivalent to $(4,3,2,1)$
-   Because of this, we proposed a new weighting function\
    $$ wI_m(r,\hat{r}) = \sum_r \sum_{\hat{r}} w(r,\hat{r}) * p(r, \hat{r}) \log{\frac{p(r,\hat{r})}{p(r)p(\hat{r})}} $$
-   The final metric we used normalizes the results: $\frac{wI_m(r,\hat{r})}{H(r,\hat{r})}$

@@

@@
